The prime of their life - the halflife of acting talent

### 2017-08-28 01_actor_careers/05_report.Rmd
Prose

Much analysis has been completed regarding the performing peak of athletes and sportspeople. Statistically, swimmers peak at around 24, marathoner runners between 30-31, gymnasts at a mere 18, NBA players generally peak between 26-29, professional hockey players between 27-28, and chess players at 30-35.

Does the idea of a performing prime extend to artists and specifically actors and directors? There are obvious reasons to be skeptical. There's no major physical component to acting that could decay as the body does, and unlike chess you wouldn't think there's the need for an exreme state of concentration and cognition which might be unreachable in older age. There are other confounding factors that might prevent us from answering this question difinitively (which I'll discuss later) but let's dig into the data.

@@

Variable | Description
-------- | -----------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seemingly ordered by screen time)

@@

# Where are the female directors?

To add a layer to this analysis I wanted group together people by gender and film production role (or actor vs director). So we think we'd be working with 4 groups, however, I was suprised to see that of our 250 top directors only XXX were female which is pretty appaling.

This got me thinking about the rest of the dataset. Given our **12,258** films dataset isn't limited to director credits in our top 250 (the film might just have a top 1000 actor appear in it) we can get a bit of a better idea of the broader gender split of directors. I've captured a lot of the key numbers in the following chart.

```{r}
```

# So

>>>>

Much analysis has been completed regarding the performing peak of athletes and sportspeople. Statistically, swimmers peak at around 24, marathon runners between 30-31, gymnasts at a mere 18, NBA players generally peak between 26-29, professional hockey players between 27-28, and chess players at 30-35.

Does the idea of a performing prime extend to artists, specifically actors and directors? There are obvious reasons to be skeptical. Importantly there's no major physical component to acting that would decay as the body does, and unlike chess you wouldn't think there's the need for an exreme state of concentration and cognition which might be unreachable at older ages.

There are significant confounding factors that might prevent us from answering this question difinitively (which I'll discuss later) but let's dig into the data.

@@

Variable | Description
------------ | ---------------------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seemingly ordered by screen time)

@@

# What's the deal with female directors?

To add a layer to this analysis I wanted group people by gender and film production role (actor vs director). So we think we'd be working with 4 groups, however, I was suprised to see that of our 250 top directors only XXX were female. A consequence of history but appaling nonetheless.

This got me thinking about the rest of the dataset. Given our **12,258** films dataset isn't limited to director credits in our top 250 (films in the set are contributed by both actor and director credits) we can get a bit of a better idea of the broader gender split of directors. I've captured a lot of the key numbers in the following chart.

```{r}
# Directors

# Top 250
top250 = films %>%
          select("name" = director) %>%
          inner_join(film_folk) %>%
          group_by(gender) %>%
          summarise(
            films = n(),
            directors = n_distinct(name)
          )

# How old are we when we produce films?
